Interval Probability for Fuzzy Quantum Theories

### 2016-05-07 prob.tex
sec 2 good more or less

% \makeatletter
% \newtheoremstyle{indented}
%   {3pt}% space before
%   {3pt}% space after
%   {\addtolength{\@totalleftmargin}{3.5em}
%    \addtolength{\linewidth}{-3.5em}
%    \parshape 1 3.5em \linewidth}% body font
%   {}% indent
%   {\bfseries}% header font
%   {.}% punctuation
%   {.5em}% after theorem header
%   {}% header specification (empty for default)
% \makeatother
% \theoremstyle{indented}
\theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{cor}{Corollary}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\pr}[2]{\langle #1, #2 \rangle}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ip}[2]{\langle #1 | #2 \rangle}
\newcommand{\proj}[1]{|#1 \rangle\langle #1 |}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\poss}{{\mbox{\wesa{possible}}}}
\newcommand{\imposs}{{\mbox{\wesa{impossible}}}}
\newcommand{\likely}{{\mbox{\wesa{likely}}}}
\newcommand{\unlikely}{{\mbox{\wesa{unlikely}}}}
\newcommand{\necess}{{\mbox{\wesa{necessary}}}}
\newcommand{\overflow}{{\mbox{\wesa{overflow}}}}

@@

\begin{document}
\title{Discrete Probability for Discrete Quantum Theories}
\author{}
\date{\today}
\maketitle

@@

@@

A \emph{probability space} specifies the necessary conditions for
reasoning coherently about collections of uncertain events.  We review
the conventional presentation of probability spaces and then discuss
several variations that avoid using the real interval $[0,1]$.

@@

\begin{example}[Two-coin probability space]\label{ex1}
  Consider an experiment that tosses two coins. We have four possible
  outcomes that constitute the sample space
  $\Omega = \{ HH, HT, TH, TT \}$. There are 16 total events including
  for example the event $\{ HH, HT \}$ that the first coin lands heads
  up, the event $\{ HT, TH \}$ that the two coins land on opposite
  sides, and the event $\{ HT, TH, TT\}$ that at least one coin lands
  tails up. Here is a possible probability measure for these events:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
  \end{array}\]
  The assignment satisfies the two constraints for probability measures:
  the probability of the entire sample space is 1, and the probability
  of every collection of disjoint events (e.g.,
  $\{ HT \} \cup \{ TH \} = \{ HT, TH \}$) is the sum of the individual
  probabilities. The probability of collections of non-disjoint events
  (e.g., $\{ HT, TH \} \cup \{ TH , TT \} = \{ HT, TH, TT \}$) may add
  to something different than the probabilities of the individual
  events. It is useful to think that this probability measure is
  completely induced by the two coins in question and their
  characteristics in the sense that each pair of coins induces a
  measure, and each measure must correspond to some pair of coins.  The
  measure above is induced by two coins such that the first coin is
  twice as likely to land tails up than heads up and the second coin is
  double-headed.  
  \qed\end{example}

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware computational or experimental setting,
it is more appropriate to consider probability measures that map
events to a finite set of elements computable with a fixed set of
resources. We expand on this observation and then consider two
approaches from the literature: set-valued probability
measures~\cite{Artstein1972,PuriRalescu1983} and interval-valued
probability
measures~\cite{Dempster1967,Zadeh1986,Weichselberger2000,JamisonLodwick2004}.

@@

Suppose we drop a needle of length $\ell$ onto a floor made of equally
spaced parallel lines a distance $\ell$ apart. It is a known fact that
the probability of the needle crossing a line is
$\frac{2}{\pi} = 0.6366197723675814$. We analyze this situation in the
mathematical framework of probability spaces paying special attention
to the resources needed to estimate the probability computationally or
experimentally.

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles and observe the number $X$ of needles
that cross a line. The sample space can be expressed as the set
$(X|-)^N$ of sequences of characters of length $N$ where each
character is either $X$ to indicate a needle crossing a line or $-$ to
indicate a needle not crossing a line. If $N=3$, the event that
exactly 2 needles cross lines is $\{-XX,X-X,XX-\}$: it has
probability~$\frac{2}{3}$. Generally, the event that exactly $M$
needles out of the $N$ total needles cross lines is $\frac{M}{N}$.

In an actual experiment with 500 needles, it was found that 321
crossed a line which gives a probability of $0.642$. In a larger
experiment with $3408$ needles, the probability was
$0.6366197193098592$. Comparing the values to the idealized
mathematical value, we see that the observed probability approaches
$\frac{2}{\pi}$ but only if \emph{larger and larger resources} are
expended.

These resource considerations thus lead us to replace the real
interval $[0,1]$ with rational numbers up to a certain precision.  The
intuition is that the precision depends on the amount of resources
allocated for the measurement process.

There is however another more subtle source of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly
$\ell$. and the needles lengths are not all equal and are not all
absolutely equal to $\ell$. These perturbations make the events
``fuzzy.'' For example, instead of talking about the idealized event
that exactly $M$ needles cross lines, we might instead talk about the
event that $M-\delta$ needles evidently cross lines and $M+\delta'$
needles plausibly cross lines. This fuzzy notion of events leads us to
consider intervals of probability reflecting the certainty of the
event and its plausibility.

%%%%%
\subsection{Set-valued Probability Measures}

Instead of using every point in the real interval $[0,1]$ as a
potential value for a probability measure, we can lump points together
in sets and consider probability measures up to set membership. The
simplest such situation is to partition the interval $[0,1]$ into two
sets: the set $\{0\}$ (which we will call $\imposs$) and the set
containing all the points in the half-open interval $(0,1]$ (which we
will call $\poss$). We will call the resulting collection
$\left\{ \imposs,\poss\right\} $, the collection $\mathscr{L}_{2}$. To
completely specify a probability space, we also need to specify how to
``add'' the probabilities in $\mathscr{L}_2$. Given two subsets of the
real interval, $I_1$ and $I_2$, the general formula for adding them
is:\footnote{Notice that the set-valued measure
  in~\cite{Artstein1972,PuriRalescu1983}, only require
  $I_{1}+I_{2}=\left\{ a_{1}+a_{2}\middle|a_{1}\in I_{1},a_{2}\in
    I_{2}\right\} $
  because they focused on measure theory. In probability theory,
  $2\in\pmeas\left(E\right)$ doesn't really make sense so that we
  consider intersect $[0,1]$ in our definition.}
\[
I_{1} \vee I_{2}=\left\{ x+y~\middle|~ x \in I_{1}, y \in I_{2}\right\} ~\cap~ [0,1]\textrm{ .}
\]
In the case of $\mathscr{L}_2$, this reduces to $x \vee y = \imposs$
iff $x = y = \imposs$.

\amr{
From Puri and Ralescu's paper: Fix $n=1$ (only 1-dimension). Conditions on $\mu$:
\begin{itemize}
\item $\mu(E) \neq \emptyset$ for every $E \in \events$, 
\item $1 \in \mu(\Omega)$, 
\item for every disjoint family $A_i$ we have, $\mu(\bigcup_i A_i) =
  \sum_i \mu(A_i)$
\end{itemize}
where the sum of subsets of $\mathbb{R}$ (not $\mathbb{R}^n$ as in the
paper) is as follows. Let $X$ and $Y$ be subsets of $\mathbb{R}$, then
$X + Y = \{ x + y ~|~ x \in X, y \in Y \}$
}

To summarize, an $\mathscr{L}$-valued probability is a function
$\pmeas:\events\rightarrow\mathscr{L}$ such that:
\begin{itemize}
\item $1\in\pmeas(\Omega)$, and 
\item for a collection $E_{i}$ of pairwise disjoint events, $\pmeas(\bigcup_{i}E_{i})=\bigvee_i\pmeas(E_{i})$. 
\end{itemize}

\begin{example}{[}Two-coin probability space with finite set-valued
probability measure{]} \label{ex2} Under the new set-valued requirement,
the probability measure in the first example becomes: 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & \imposs\\
\pmeas(\{HH\}) & = & \poss\\
\pmeas(\{HT\}) & = & \imposs\\
\pmeas(\{TH\}) & = & \poss\\
\pmeas(\{TT\}) & = & \imposs\\
\pmeas(\{HH,HT\}) & = & \poss\\
\pmeas(\{HH,TH\}) & = & \poss\\
\pmeas(\{HH,TT\}) & = & \poss
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & \poss\\
\pmeas(\{HT,TT\}) & = & \imposs\\
\pmeas(\{TH,TT\}) & = & \poss\\
\pmeas(\{HH,HT,TH\}) & = & \poss\\
\pmeas(\{HH,HT,TT\}) & = & \poss\\
\pmeas(\{HH,TH,TT\}) & = & \poss\\
\pmeas(\{HT,TH,TT\}) & = & \poss\\
\pmeas(\{HH,HT,TH,TT\}) & = & \poss
\end{array}\end{array}
\]
Despite the fact that we have lost all numeric information, the probability
measure still reveals that the second coin is double-headed. We have
however lost the information regarding the bias in the first coin.
This information can be recovered with a more refined probability
measure as we show next. \qed\end{example}

Although $\mathscr{L}_{2}$-valued probability measure is quite intuitive,
set-valued probability measure is not that intuitive if we extend
to more values. For example, consider the following three \emph{overlapping}
closed sub-intervals: $[0,0]$ is still call \imposs, $[0,\frac{1}{2}]$
which we call \unlikely, ad $[\frac{1}{2},1]$ which we call \likely.
The following example is a set-valued probability measure corresponding
to the probability measure in the first example.

\begin{example}\label{Two-coin-set-valued-again}{[}Two-coin probability
space with set-valued probability measure, again{]} 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & \imposs\\
\pmeas(\{HH\}) & = & \unlikely\\
\pmeas(\{HT\}) & = & \imposs\\
\pmeas(\{TH\}) & = & \likely\\
\pmeas(\{TT\}) & = & \imposs\\
\pmeas(\{HH,HT\}) & = & \unlikely\\
\pmeas(\{HH,TH\}) & = & \likely\\
\pmeas(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & \likely\\
\pmeas(\{HT,TT\}) & = & \imposs\\
\pmeas(\{TH,TT\}) & = & \likely\\
\pmeas(\{HH,HT,TH\}) & = & \likely\\
\pmeas(\{HH,HT,TT\}) & = & \unlikely\\
\pmeas(\{HH,TH,TT\}) & = & \likely\\
\pmeas(\{HT,TH,TT\}) & = & \likely\\
\pmeas(\{HH,HT,TH,TT\}) & = & \likely
\end{array}\end{array}
\]
In this example, we can get the information that the first coin is
weighted and the second coin is double-headed. However, we may notice
that the set-valued probability of the whole space~$\{HH,HT,TH,TT\}$
are different between these two examples. This problem can be fixed
when we moved on to the next section. \qed\end{example}

We will return to finite set-valued probability measures in Sec.~\ref{sec:?}.

%%%%%

\subsection{Interval-valued probability measures}

The singleton closed interval $[0,0]$ is a required spacial value
for empty set, and a natural generalization is to consider another
special closed interval $[1,1]$ which we call \necess for the whole
space. In general, if the probability of an event $E$ is $[a,b]$,
we think of the left-endpoint~$a$ as representing the strength of
the evidence that supports $E$, and the right-endpoint~$b$ as the
strength of the evidence that contradicts $E$.

Thus if we have an event $E$ with probability $[a,b]$ where $a=0.1$
and $b=0.7$, we have that: 
\begin{itemize}
\item the strength of evidence supporting $E$ is 0.1; since either $E$
or its complement must happen, we conclude that there is 0.9 evidence
supporting the complement of $E$; 
\item the strength of evidence contradicting $E$ is 0.7; again since either
$E$ or its complement must happen, we conclude that there is 0.3
evidence contradicting the complement of $E$. 
\end{itemize}
\yutsung{Do we use the law of excluded middle here? You remind me
Agda : ) Did Homotopy Type Theory people said anything about the probability?}\\
Turning things around, the strength of evidence that contradicts $E$
is evidence supporting the complement of $E$. The complement of $E$
must therefore have probability $\left[1-b,1-a\right]$ which we abbreviate
$1-\left[a,b\right]$, so the probability measure $\pmeas$ should
satisfy:
\begin{itemize}
\item $\pmeas(\emptyset)=[0,0]$, 
\item $\pmeas(\Omega)=[1,1]$, and 
\item $\pmeas\left(\Omega\backslash E\right)=1-\pmeas\left(E\right)$
\end{itemize}
However, if we want $\pmeas(\Omega)=[1,1]$, and the probability assignment
for singleton sets $\pmeas(\{HH\})=\unlikely$, $\pmeas(\{HT\})=\imposs$,
$\pmeas(\{TH\})=\likely$, and $\pmeas(\{TT\})=\imposs$ as in example~\ref{Two-coin-set-valued-again},
we only have $\pmeas(\Omega)\subsetneq\pmeas(\{HH\})+\pmeas(\{HT\})+\pmeas(\{TH\})+\pmeas(\{TT\})$.
In general, we only require
\begin{itemize}
\item for a collection $E_{i}$ of pairwise disjoint events, we have $\pmeas\left(\bigcup_{i}E_{i}\right)\subseteq\sum_{i}\pmeas\left(E_{i}\right)$. 
\end{itemize}
This statement means that the evidences of $\bigcup_{i}E_{i}$ is
at least as strong as putting all the evidences of $E_{i}$ together,
but some evidence may only be acquired for $\bigcup_{i}E_{i}$ as
the whole. Therefore, $\pmeas\left(\bigcup_{i}E_{i}\right)$ is a
subset of $\sum_{i}\pmeas\left(E_{i}\right)$, but may not equal.
In our example, 
\begin{eqnarray*}
 &  & \pmeas(\{HH\})+\pmeas(\{HT\})+\pmeas(\{TH\})+\pmeas(\{TT\})\\
 & = & \imposs+\unlikely+\imposs+\likely\\
 & = & \left[0,0\right]+\left[0,\frac{1}{2}\right]+\left[0,0\right]+\left[\frac{1}{2},1\right]=\left[\frac{1}{2},\frac{3}{2}\right]
\end{eqnarray*}
The above equation told us $\pmeas(\Omega)\subseteq\left[\frac{1}{2},\frac{3}{2}\right]$.
However, because of $\pmeas(\emptyset)=\imposs$, we have $\pmeas\left(\Omega\right)=1-\pmeas\left(\emptyset\right)=\necess$,
but $\pmeas(\emptyset)=\imposs$ cannot be used to reasoning the probability
of $\{HH\}$, $\{HT\}$, $\{TH\}$ and $\{TT\}$ individually. The
full probability assignment is shown in the following example.

>>>>

% \makeatletter
% \newtheoremstyle{indented}
%   {3pt}% space before
%   {3pt}% space after
%   {\addtolength{\@totalleftmargin}{3.5em}
%    \addtolength{\linewidth}{-3.5em}
%    \parshape 1 3.5em \linewidth}% body font
%   {}% indent
%   {\bfseries}% header font
%   {.}% punctuation
%   {.5em}% after theorem header
%   {}% header specification (empty for default)
% \makeatother
% \theoremstyle{indented}
\theoremstyle{remark}
\newtheorem{example}{Example}
\newtheorem{definition}{Definition}
\newtheorem{cor}{Corollary}
\newcommand{\events}{\ensuremath{\mathcal{E}}}
\newcommand{\qevents}{\ensuremath{\mathcal{E}}}
\newcommand{\pmeas}{\ensuremath{\mu}}
\newcommand{\Hilb}{\mathcal{H}}
\newcommand{\pr}[2]{\langle #1, #2 \rangle}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\ip}[2]{\langle #1 | #2 \rangle}
\newcommand{\proj}[1]{|#1 \rangle\langle #1 |}
\newcommand{\ps}{\texttt{+}}
\newcommand{\ms}{\texttt{-}}
\newcommand{\poss}{{\mbox{\wesa{possible}}}}
\newcommand{\imposs}{{\mbox{\wesa{impossible}}}}
\newcommand{\likely}{{\mbox{\wesa{likely}}}}
\newcommand{\unlikely}{{\mbox{\wesa{unlikely}}}}
\newcommand{\necess}{{\mbox{\wesa{certain}}}}
\newcommand{\overflow}{{\mbox{\wesa{overflow}}}}

@@

\begin{document}
\title{Interval Probability for Fuzzy Quantum Theories}
\author{}
\date{\today}
\maketitle

@@

Fuzzy quantum mechanics:
\begin{itemize}
\item \url{http://cds.cern.ch/record/518511/files/0107054.pdf}
\item \url{http://link.springer.com/chapter/10.1007%2F978-3-642-35644-5_18#page-1}
\item \url{http://link.springer.com/chapter/10.1007%2F978-3-540-93802-6_20#page-1}
\item \url{http://www.du.edu/nsm/departments/mathematics/media/documents/preprints/m0412.pdf}
\item \url{http://www.space-lab.ru/files/pages/PIRT_VII-XII/pages/text/PIRT_X/Bobola.pdf}
\item \url{http://www.vub.ac.be/CLEA/aerts/publications/1993LiptovskyJan.pdf}
\end{itemize}

\noindent Pseudo-randomness:
\begin{itemize}
\item
  \url{https://people.csail.mit.edu/silvio/Selected%20Scientific%20Papers/Pseudo%20Randomness/How_To_Generate_Cryptographically_Strong_Sequences_Of_Pseudo-Random_Bits.pdf}:
  ``the randomness of an event is relative to a specific model of
  computation with a specified amount of computing resources.''
\item Another version \url{https://pdfs.semanticscholar.org/3e9c/5f6f48d9ef426655dc799e9b287d754e86c1.pdf}
\end{itemize}

@@

A \emph{probability space} specifies the necessary conditions for
reasoning coherently about collections of uncertain events.  We review
the conventional presentation of probability spaces and then discuss
the hidden issue of computational and experimental resources.

@@

\begin{example}[Two-coins experiment]\label{ex1}
  Consider an experiment that tosses two coins. We have four possible
  outcomes that constitute the sample space
  $\Omega = \{ HH, HT, TH, TT \}$. There are 16 total events including
  for example the event $\{ HH, HT \}$ that the first coin lands heads
  up, the event $\{ HT, TH \}$ that the two coins land on opposite
  sides, and the event $\{ HT, TH, TT\}$ that at least one coin lands
  tails up. Here is a possible probability measure for these events:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
  \end{array}\]
  The assignment satisfies the two constraints for probability measures:
  the probability of the entire sample space is 1, and the probability
  of every collection of disjoint events (e.g.,
  $\{ HT \} \cup \{ TH \} = \{ HT, TH \}$) is the sum of the individual
  probabilities. The probability of collections of non-disjoint events
  (e.g., $\{ HT, TH \} \cup \{ TH , TT \} = \{ HT, TH, TT \}$) may add
  to something different than the probabilities of the individual
  events. It is useful to think that this probability measure is
  completely induced by the two coins in question and their
  characteristics in the sense that each pair of coins induces a
  measure, and each measure must correspond to some pair of coins.  The
  measure above is induced by two coins such that the first coin is
  twice as likely to land tails up than heads up and the second coin is
  double-headed.  
  \qed\end{example}

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware computational or experimental setting,
it is more appropriate to consider probability measures that map
events to a finite set of elements computable with a fixed set of
resources. We expand on this observation and then consider 
%% two approaches from the literature: set-valued probability
%% measures~\cite{Artstein1972,PuriRalescu1983} and 
interval-valued probability
measures~\cite{Dempster1967,Zadeh1986,Weichselberger2000,JamisonLodwick2004}.

@@

Suppose we drop a needle of length $\ell$ onto a floor made of equally
spaced parallel lines a distance $\ell$ apart. It is a known fact that
the probability of the needle crossing a line converges to
$\frac{2}{\pi} = 0.6366197723675814$.
%% \yutsung{Add $\ldots$ at the end? }\\
We analyze this situation in the mathematical framework of probability
spaces paying special attention to the resources needed to estimate
the probability computationally or experimentally.

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles and observe the number $X$ of needles
that cross a line. The sample space can be expressed as the set
$(X|-)^N$
%% \yutsung{Use more physical or mathematical notation?}\\
of sequences of characters of length $N$ where each character is
either $X$ to indicate a needle crossing a line or $-$ to indicate a
needle not crossing a line. If $N=3$, the event that exactly 2 needles
cross lines is $\{-XX,X-X,XX-\}$: it has
probability~$\frac{2}{3}$. Generally, the event that exactly $M$
needles out of the $N$ total needles cross lines is $\frac{M}{N}$.

In an actual experiment with 500 needles, it was found that 321
crossed a line which gives a probability of $0.642$. In a larger
experiment with $3408$ needles, the probability was calculated to be
$0.6366197193098592$.
%% \yutsung{Add $\ldots$ at the end? }\\
Comparing the values to the idealized mathematical value, we see that
the observed probability approaches $\frac{2}{\pi}$ but only if
\emph{larger and larger resources} are expended.

These resource considerations thus lead us to replace the real
interval $[0,1]$ with rational numbers up to a certain precision.  The
intuition is that the precision depends on the amount of resources
allocated for the measurement process. There is clearly a connection
between the number of needles and the achievable precision: in the
hypothetical experiment with 3 needles, it is not sensible to retain
100 digits in the expansion of $\frac{2}{3}$.

There is however another more subtle source of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly
$\ell$. and the needles lengths are not all equal and are not all
absolutely equal to $\ell$. These perturbations make the events
``fuzzy.'' For example, instead of talking about the idealized event
that exactly $M$ needles cross lines which would require the most
expensive needles built to the most precise accuracy, laser precision
for drawing lines on the floor, and the most powerful microscopes to
determine if a needle does cross a line, we might instead talk about
the event that $M-\delta$ needles evidently cross lines and
$M+\delta'$ needles plausibly cross lines for some acceptable
approximations $\delta$ and $\delta'$. This fuzzy notion of events
leads us to consider intervals of probability reflecting the certainty
of the event and its plausibility. In the hypothetical experiment with
3 needles, assume that one needle clearly crossed a line and another
was close enough to plausibly be considered as crossing the line: we
would express the probability in this case as the interval
$[\frac{1}{3},\frac{2}{3}]$ expressing that we are certain that the
event has probability at least $\frac{1}{3}$ but it is possible that
it would have probability $\frac{2}{3}$.

% %%%%%
% \subsection{Set-valued Probability Measures}

% Instead of using every point in the real interval $[0,1]$ as a
% potential value for a probability measure, we can lump points together
% in sets and consider probability measures up to set membership. The
% simplest such situation is to partition the interval $[0,1]$ into two
% sets: the set $\{0\}$ (which we will call $\imposs$) and the set
% containing all the points in the half-open interval $(0,1]$ (which we
% will call $\poss$). We will call the resulting collection
% $\left\{ \imposs,\poss\right\} $, the collection $\mathscr{L}_{2}$. To
% completely specify a probability space, we also need to specify how to
% ``add'' the probabilities in $\mathscr{L}_2$. Given two subsets of the
% real interval, $I_1$ and $I_2$, the general formula for adding them
% is:\footnote{Notice that the set-valued measure
%   in~\cite{Artstein1972,PuriRalescu1983}, only require
%   $I_{1}+I_{2}=\left\{ a_{1}+a_{2}\middle|a_{1}\in I_{1},a_{2}\in
%     I_{2}\right\} $
%   because they focused on measure theory. In probability theory,
%   $2\in\pmeas\left(E\right)$ doesn't really make sense so that we
%   consider intersect $[0,1]$ in our definition.}
% \[
% I_{1} \vee I_{2}=\left\{ x+y~\middle|~ x \in I_{1}, y \in I_{2}\right\} ~\cap~ [0,1]\textrm{ .}
% \]
% In the case of $\mathscr{L}_2$, this reduces to $x \vee y = \imposs$
% iff $x = y = \imposs$.

% \amr{
% From Puri and Ralescu's paper: Fix $n=1$ (only 1-dimension). Conditions on $\mu$:
% \begin{itemize}
% \item $\mu(E) \neq \emptyset$ for every $E \in \events$, 
% \item $1 \in \mu(\Omega)$, 
% \item for every disjoint family $A_i$ we have, $\mu(\bigcup_i A_i) =
%   \sum_i \mu(A_i)$
% \end{itemize}
% where the sum of subsets of $\mathbb{R}$ (not $\mathbb{R}^n$ as in the
% paper) is as follows. Let $X$ and $Y$ be subsets of $\mathbb{R}$, then
% $X + Y = \{ x + y ~|~ x \in X, y \in Y \}$
% }\\
% \yutsung{Actually, using $X+Y=\{x+y~|~x\in X,y\in Y\}$ directly
% might be more confusing than I thought... Because our example will
% be follows:

% For $\{0\}=\imposs$ and $(0,\infty)=\poss$, it will become
% \[
% \begin{array}{c@{\qquad\qquad}c}
% \begin{array}{rcl}
% \pmeas(\emptyset) & = & \{0\}\\
% \pmeas(\{HH\}) & = & (0,\infty)\\
% \pmeas(\{HT\}) & = & \{0\}\\
% \pmeas(\{TH\}) & = & (0,\infty)\\
% \pmeas(\{TT\}) & = & \{0\}\\
% \pmeas(\{HH,HT\}) & = & (0,\infty)\\
% \pmeas(\{HH,TH\}) & = & (0,\infty)\\
% \pmeas(\{HH,TT\}) & = & (0,\infty)
% \end{array} & \begin{array}{rcl}
% \pmeas(\{HT,TH\}) & = & (0,\infty)\\
% \pmeas(\{HT,TT\}) & = & \{0\}\\
% \pmeas(\{TH,TT\}) & = & (0,\infty)\\
% \pmeas(\{HH,HT,TH\}) & = & (0,\infty)\\
% \pmeas(\{HH,HT,TT\}) & = & (0,\infty)\\
% \pmeas(\{HH,TH,TT\}) & = & (0,\infty)\\
% \pmeas(\{HT,TH,TT\}) & = & (0,\infty)\\
% \pmeas(\{HH,HT,TH,TT\}) & = & (0,\infty)
% \end{array}\end{array}
% \]

% For $\{0\}=\imposs$ and $(0,1]=\poss$, it will become the following
% and gives an extra-value~$(0,2]$.
% \[
% \begin{array}{c@{\qquad\qquad}c}
% \begin{array}{rcl}
% \pmeas(\emptyset) & = & \{0\}\\
% \pmeas(\{HH\}) & = & (0,1]\\
% \pmeas(\{HT\}) & = & \{0\}\\
% \pmeas(\{TH\}) & = & (0,1]\\
% \pmeas(\{TT\}) & = & \{0\}\\
% \pmeas(\{HH,HT\}) & = & (0,1]\\
% \pmeas(\{HH,TH\}) & = & (0,2]\\
% \pmeas(\{HH,TT\}) & = & (0,1]
% \end{array} & \begin{array}{rcl}
% \pmeas(\{HT,TH\}) & = & (0,1]\\
% \pmeas(\{HT,TT\}) & = & \{0\}\\
% \pmeas(\{TH,TT\}) & = & (0,1]\\
% \pmeas(\{HH,HT,TH\}) & = & (0,2]\\
% \pmeas(\{HH,HT,TT\}) & = & (0,1]\\
% \pmeas(\{HH,TH,TT\}) & = & (0,2]\\
% \pmeas(\{HT,TH,TT\}) & = & (0,1]\\
% \pmeas(\{HH,HT,TH,TT\}) & = & (0,2]
% \end{array}\end{array}
% \]

% For $\{0\}=\imposs$, $[0,\frac{1}{2}]=\unlikely$, and $[\frac{1}{2},1]=\likely$,
% it will become the following and gives an extra-value~$[\frac{1}{2},\frac{3}{2}]$.
% \[
% \begin{array}{c@{\qquad\qquad}c}
% \begin{array}{rcl}
% \pmeas(\emptyset) & = & \{0\}\\
% \pmeas(\{HH\}) & = & [0,\frac{1}{2}]\\
% \pmeas(\{HT\}) & = & \{0\}\\
% \pmeas(\{TH\}) & = & [\frac{1}{2},1]\\
% \pmeas(\{TT\}) & = & \{0\}\\
% \pmeas(\{HH,HT\}) & = & [0,\frac{1}{2}]\\
% \pmeas(\{HH,TH\}) & = & [\frac{1}{2},\frac{3}{2}]\\
% \pmeas(\{HH,TT\}) & = & [0,\frac{1}{2}]
% \end{array} & \begin{array}{rcl}
% \pmeas(\{HT,TH\}) & = & [\frac{1}{2},1]\\
% \pmeas(\{HT,TT\}) & = & \{0\}\\
% \pmeas(\{TH,TT\}) & = & [\frac{1}{2},1]\\
% \pmeas(\{HH,HT,TH\}) & = & [\frac{1}{2},\frac{3}{2}]\\
% \pmeas(\{HH,HT,TT\}) & = & [0,\frac{1}{2}]\\
% \pmeas(\{HH,TH,TT\}) & = & [\frac{1}{2},\frac{3}{2}]\\
% \pmeas(\{HT,TH,TT\}) & = & [\frac{1}{2},1]\\
% \pmeas(\{HH,HT,TH,TT\}) & = & [\frac{1}{2},\frac{3}{2}]
% \end{array}\end{array}
% \]

% Actually, this approach is more natural to me when I think in measure
% theory, but I guess the readers might be confused if write in this
% way?}

% To summarize, an $\mathscr{L}$-valued probability is a function
% $\pmeas:\events\rightarrow\mathscr{L}$ such that:
% \begin{itemize}
% \item $1\in\pmeas(\Omega)$, and 
% \item for a collection $E_{i}$ of pairwise disjoint events, $\pmeas(\bigcup_{i}E_{i})=\bigvee_i\pmeas(E_{i})$. 
% \end{itemize}
